---
title: "Plotting simmer statistics"
author: "Bart Smeets, Iñaki Ucar"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Plotting simmer statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>", 
                      fig.width = 6, fig.height = 4, fig.align = "center")
```

## Simulation example

This is the basic example from [_Introduction to simmer_](https://CRAN.R-project.org/package=simmer/vignettes/A-introduction.html). Please, refer to that vignette for further detail.

```{r, message=FALSE}
library(simmer)
library(parallel)

patient <- trajectory("patients' path") %>%
  ## add an intake activity 
  seize("nurse", 1) %>%
  timeout(function() rnorm(1, 15)) %>%
  release("nurse", 1) %>%
  ## add a consultation activity
  seize("doctor", 1) %>%
  timeout(function() rnorm(1, 20)) %>%
  release("doctor", 1) %>%
  ## add a planning activity
  seize("administration", 1) %>%
  timeout(function() rnorm(1, 5)) %>%
  release("administration", 1)

envs <- mclapply(1:100, function(i) {
  simmer("SuperDuperSim") %>%
    add_resource("nurse", 1) %>%
    add_resource("doctor", 2) %>%
    add_resource("administration", 1) %>%
    add_generator("patient", patient, function() rnorm(1, 10, 2)) %>%
    run(80) %>%
    wrap()
})
```

## Basic visualisation tools

This package provides some basic visualisation tools to help you take a glance at your simulations quickly, but of course you probably may want to extract the raw data and develop your own data analysis.

There are three types of plot implemented with different metrics available:

- Plot of resources. Two metrics available:
    - the usage of a resource over the simulation time frame.
    - the utilization of specified resources in the simulation.
- Plot of arrivals. Three metrics available:
    - activity time.
    - waiting time.
    - flow time.
- Plot of attributes.

All these plots and metrics are accessible through the S3 method for the `plot` generic provided by this package (see `?plot.simmer` for more details), which uses the `ggplot2` package as backend.

For instance, after you've left it simmering for a bit (pun intended), we can have a look at the overall resource utilization. The top and bottom of the error bars show respectively the 25th and 75th percentile of the utilization across all the replications. The top of the bar shows the median utilization.

```{r, message=FALSE}
library(simmer.plot)

plot(envs, what = "resources", metric = "utilization", c("nurse", "doctor","administration"))
```

It is also possible to have a look at resources' activity during the simulation.

```{r, message=FALSE}
plot(envs, what = "resources", metric = "usage", c("nurse", "doctor"), items = "server", steps = T)
```

In the above graph, the individual lines are all separate replications. The step lines are instantaneous utilization and the smooth line is a running average. Let's take a look now at a specific replication for one resource. In the example below the 6th replication is shown.

```{r, message=FALSE}
plot(envs[[6]], what = "resources", metric = "usage", "doctor", items = "server", steps = T)
```

Next we can have a look at the evolution of the arrivals' flow time during the simulation. In the plot below, each individual line represents a replication. A smooth line is drawn over them. All arrivals that didn't finish their entire trajectory are excluded from the plot.

```{r, message=FALSE}
plot(envs, what = "arrivals", metric = "flow_time")
```

Similarly one can have a look at the evolution of the activity times with ```metric = "activity_time"``` and waiting times with ```metric = "waiting_time"```.
